Bayesian analysis of the NESTA study of interventions against  verbal aggression online

### 2021-09-16 bayesianAnalysis/bayesianReport3.Rmd
direct effect

The situation is different when it comes to evaluating the *direct* effect of intervention. Then, we also need to block indirect causal paths from the intervention to the outcome.

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
adjustmentSets(dag, exposure = c("IC", "IT"), outcome = "AAS", effect = "direct")
```
\normalsize

For such an evaluation we need to also condition on \textsf{CAS}, which is what we will do when we turn to the study of the direct effects of the interventions.

@@

# Bayesian models, priors and diagnostics

@@

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", results= "hide"}
# building model with sd=1
# InteractionsModelDiffSD1 <- ulam(
#   alist(
#     AdiffS ~ dnorm( mu, sigma ),
#     mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC+
#     bADSIC * ADS * IC+ bCBS[groupID] *CBS,
#     a ~ dnorm (0,1),
#     bADS[groupID] ~ dnorm(0,1),
#     bADSIC ~ dnorm(0,1),
#     bCBS[groupID] ~ dnorm(0,1),
#     bIT[groupID] ~ dnorm(0,1),
#     bIC[groupID] ~ dnorm(0,1),
#     sigma  ~ dexp(1)
#   ), 
#   data = summaries
# )
# 
# saveRDS(InteractionsModelDiffSD1, file = "models/InteractionsModelDiffSD1.rds")
InteractionsModelDiffSD1 <- readRDS(file = "models/InteractionsModelDiffSD1.rds")

@@

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", results= "hide"}
#prior predictive check sd =.3
ADS <- 0
CBS <- 0
groupID <- 1:3
IC <- 5  #mean for interventions in treatment
data <- expand.grid(ADS = ADS,groupID = groupID, CBS = CBS, IC =  IC)
prior <- extract.prior(InteractionsModelDiff, n = 1e4)
mu <- link(InteractionsModelDiff , post=prior , data=data ) 
colnames(mu) <- levels(summaries$group)
muLong <- melt(mu)
colnames(muLong) <- c("id", "group", "AdiffS")
head(muLong)

@@

<!-- # InteractionsModelDiff <- ulam( -->
<!-- #   alist( -->
<!-- #     AdiffS ~ dnorm( mu, sigma ), -->
<!-- #     mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + bADSIC * ADS * IC+ bCBS[groupID] *CBS, -->
<!-- #     a ~ dnorm (0,0.3), -->
<!-- #     bADS[groupID] ~ dnorm(0,.3), -->
<!-- #     bADSIC ~ dnorm(0,.3), -->
<!-- #     bCBS[groupID] ~ dnorm(0,.3), -->
<!-- #     bIT[groupID] ~ dnorm(0,.3), -->
<!-- #     bIC[groupID] ~ dnorm(0,.3), -->
<!-- #     sigma  ~ dexp(1) -->
<!-- #   ),  -->
<!-- #   data = summaries -->
<!-- # ) -->

@@

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
comparison<- compare(null,ADS,ADSIC,IT,ADSIT,ADSITIC,ADSITIC_ADSIC,
                     ADSITIC_ADSIC_ADSIT,ADSIT_ADSIT,ADSITIC_ADSIT_ITIC_ADSIC,
                     ADSITICCBS_ITIC_ADSIC,Final, tooFar)
mykable(data.frame(comparison )) 
plot(comparison)
```
\normalsize

@@

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
precis(FinalHMC, depth = 2)
plot(precis(FinalHMC, depth = 2))
```
\normalsize

@@

  return(ggplot(visContrastTreatment, aes(x = CBS, y = contrast, color = group ))+
            geom_pointrange(mapping = 
        aes(ymin = cLow, ymax = cHigh), size = .05, 
        fatten = .1, alpha = .5,
        position = position_dodge(width = .8))+
          theme_tufte())
}

visContrastCBSJoint <- ggarrange(visContrastsCBS(FinalHMC,ADS = -2)+
      ggtitle("ADS = -2")+ylim(c(-2.5,2.5))+ scale_color_discrete(guide=FALSE),
          visContrastsCBS(FinalHMC,ADS = 0)+ggtitle("ADS = 0")+
        ylim(c(-2.5,2.5))+ scale_color_discrete(guide=FALSE)+
        removeY,
        visContrastsCBS(FinalHMC,ADS = 2)+ggtitle("ADS = 2")+
        ylim(c(-2.5,2.5))+ theme(legend.position = c(0.75, 0.15))+
        removeY, ncol = 3)

@@

  return(ggplot(visContrastTreatmentADS, aes(x = ADS, y = contrast, color = group ))+
  geom_pointrange(mapping = aes(ymin = cLow, ymax = cHigh), size = .05,
                  fatten = .1, alpha = .5, 
                  position = position_dodge(width = .8)) +theme_tufte())
}

visContrastADSJoint <- ggarrange(visContrastsADS(FinalHMC,CBS = -2)+ggtitle("CBS = -2")
                                 +ylim(c(-2.5,2.5))+ scale_color_discrete(guide=FALSE),
                                 visContrastsADS(FinalHMC,CBS = 0)+ggtitle("CBS = 0")
                                 +ylim(c(-2.5,2.5))+ scale_color_discrete(guide=FALSE)+
                                   removeY,
                                 visContrastsADS(FinalHMC,CBS = 2)+ggtitle("CBS = 2")
                                 +ylim(c(-2.5,2.5))+
                                   theme(legend.position = c(0.75, 0.15))+
                                   removeY, ncol = 3)

@@

Now, let's inspect the impact of intervention counts by treatment type by looking at contrasts (distances from the control group mean) with 89% HPDIs by IC.

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
visContrastsIC <- function(model = FinalHMC, CBS = CBS , IC =  seq(0,30,by = 1), ADS = ADS)
{
  groupID <- 1:3
  data <- expand.grid(CBS, groupID, IC , ADS)
  data
  colnames(data) <- c("CBS", "groupID", "IC", "ADS")
  posterior <- extract.samples(model, n = 1e5)
  mu <- link( model, data=data ) 
  means <-  round(apply(mu , 2 , mean ), 4)
  HPDIs <- round(apply( mu , 2 , HPDI ),4)
  visContrastIC <- cbind(data,means,t(as.data.frame(HPDIs)))

@@

  return(ggplot(visContrastTreatmentIC, aes(x = IC, y = contrast, color = group ))+
    geom_pointrange(mapping = aes(ymin = cLow, ymax = cHigh), size = .2, alpha = .5)+
    ylim(c(-2,2)) +theme_tufte())
}

visContrastsICJoint <- ggarrange(
visContrastsIC(ADS = 2, CBS = -2)+removeX+ 
               theme(legend.position = c(0.3, 0.9),
                     legend.key.size = unit(.3, 'cm'),
                     legend.key.height = unit(.3, 'cm'),
                     legend.key.width = unit(.3, 'cm'),
                     legend.title= element_blank())+
  ggtitle("CBS = -2")+
  ylab("ADS = 2"),
    visContrastsIC(ADS = 2, CBS = 0)+removeY+removeX+ scale_color_discrete(guide=FALSE)+
  ggtitle("CBS = 0"),
    visContrastsIC(ADS = 2, CBS = 2)+removeY+removeX+ggtitle("CBS = 2")+ scale_color_discrete(guide=FALSE),
visContrastsIC(ADS = 0, CBS = -2)+removeX+ scale_color_discrete(guide=FALSE)+
  ylab("ADS = 0"),
    visContrastsIC(ADS = 0, CBS = 0)+removeY+removeX+ 
  scale_color_discrete(guide=FALSE),
    visContrastsIC(ADS = 0, CBS = 2)+removeY+removeX+ scale_color_discrete(guide=FALSE),  
visContrastsIC(ADS = -2, CBS = -2)+ scale_color_discrete(guide=FALSE)+
  ylab("ADS = -2"),
    visContrastsIC(ADS = -2, CBS = 0)+removeY+ scale_color_discrete(guide=FALSE),
    visContrastsIC(ADS = -2, CBS = 0)+removeY+ scale_color_discrete(guide=FALSE), 
ncol = 3, nrow = 3
)

@@

>>>>

The situation is different when it comes to evaluating the *direct* effect of intervention. Then, we also need to block indirect causal paths from the intervention to the outcome. For such an evaluation we need to also condition on \textsf{CAS}, which is what we will do when we turn to the study of the direct effects of the interventions.

@@

@@

\vspace{1mm}
\footnotesize
```{r priors1,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", results= "hide"}
# building model with sd=1
# InteractionsModelDiffSD1 <- ulam(
#   alist(
#     AdiffS ~ dnorm( mu, sigma ),
#     mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC+
#     bADSIC * ADS * IC+ bCBS[groupID] *CBS,
#     a ~ dnorm (0,1),
#     bADS[groupID] ~ dnorm(0,1),
#     bADSIC ~ dnorm(0,1),
#     bCBS[groupID] ~ dnorm(0,1),
#     bIT[groupID] ~ dnorm(0,1),
#     bIC[groupID] ~ dnorm(0,1),
#     sigma  ~ dexp(1)
#   ), 
#   data = summaries
# )
# 
# saveRDS(InteractionsModelDiffSD1, file = "models/InteractionsModelDiffSD1.rds")
InteractionsModelDiffSD1 <- readRDS(file = "models/InteractionsModelDiffSD1.rds")

@@

\vspace{1mm}
\footnotesize
```{r priors03,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", results= "hide"}
#prior predictive check sd =.3
ADS <- 0
CBS <- 0
groupID <- 1:3
IC <- 5  #mean for interventions in treatment
data <- expand.grid(ADS = ADS,groupID = groupID, CBS = CBS, IC =  IC)
prior <- extract.prior(InteractionsModelDiff, n = 1e4)
mu <- link(InteractionsModelDiff , post=prior , data=data ) 
colnames(mu) <- levels(summaries$group)
muLong <- melt(mu)
colnames(muLong) <- c("id", "group", "AdiffS")
head(muLong)

@@

@@

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
comparison<- compare(null,ADS,ADSIC,IT,ADSIT,ADSITIC,ADSITIC_ADSIC,
                     ADSITIC_ADSIC_ADSIT,ADSIT_ADSIT,ADSITIC_ADSIT_ITIC_ADSIC,
                     ADSITICCBS_ITIC_ADSIC,Final, tooFar)
mykable(round(data.frame(comparison ),3)) 
plot(comparison)
```
\normalsize

@@

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
mykable(data.frame(precis(FinalHMC, depth = 2)))
plot(precis(FinalHMC, depth = 2))
```
\normalsize

@@

  return(ggplot(visContrastTreatment, aes(x = CBS, y = contrast, fill = group ))+
           geom_line(se = FALSE)+
            geom_ribbon(mapping = 
        aes(ymin = cLow, ymax = cHigh),  
         alpha = .3)+
          theme_tufte())
}

visContrastCBSJoint <- ggarrange(visContrastsCBS(FinalHMC,ADS = -2)+
      ggtitle("ADS = -2")+ylim(c(-2.5,2.5))+ scale_fill_discrete(guide=FALSE),
          visContrastsCBS(FinalHMC,ADS = 0)+ggtitle("ADS = 0")+
        ylim(c(-2.5,2.5))+ scale_fill_discrete(guide=FALSE)+
        removeY,
        visContrastsCBS(FinalHMC,ADS = 2)+ggtitle("ADS = 2")+
        ylim(c(-2.5,2.5))+ theme(legend.position = c(0.75, 0.15))+
        removeY, ncol = 3)

@@

  return(ggplot(visContrastTreatmentADS, aes(x = ADS, y = contrast, fill = group ))+
           geom_line(se = FALSE) +
  geom_ribbon(mapping = aes(ymin = cLow, ymax = cHigh), 
       alpha = .3) +theme_tufte())
}

visContrastADSJoint <- ggarrange(visContrastsADS(FinalHMC,CBS = -2)+ggtitle("CBS = -2")
                                 +ylim(c(-2.5,2.5))+ scale_fill_discrete(guide=FALSE),
                                 visContrastsADS(FinalHMC,CBS = 0)+ggtitle("CBS = 0")
                                 +ylim(c(-2.5,2.5))+ scale_fill_discrete(guide=FALSE)+
                                   removeY,
                                 visContrastsADS(FinalHMC,CBS = 2)+ggtitle("CBS = 2")
                                 +ylim(c(-2.5,2.5))+
                                   theme(legend.position = c(0.75, 0.15))+
                                   removeY, ncol = 3)

@@

Now, let's inspect the impact of intervention counts by treatment type by looking at contrasts (distances from the control group mean) with 89% HPDIs by IC. Notice the predicted effect of \textsf{IC} is weaker than group membership, so for visibility the $y$-axis has a smaller range. Also, not enough data was available to reliably estimate uncertainty for \textsf{IC} above 20, hence the restriction on the $x$-axis (arleady at lower values, lack of estimates is visible for the more extreme  covariate settings).

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
visContrastsIC <- function(model = FinalHMC, CBS = CBS ,
                           IC =  seq(0,20,by = 1), ADS = ADS)
{
  groupID <- 1:3
  data <- expand.grid(CBS, groupID, IC , ADS)
  data
  colnames(data) <- c("CBS", "groupID", "IC", "ADS")
  posterior <- extract.samples(model, n = 1e5)
  mu <- link( model, data=data ) 
  means <-  round(apply(mu , 2 , mean ), 4)
  HPDIs <- round(apply( mu , 2 , HPDI ),4)
  visContrastIC <- cbind(data,means,t(as.data.frame(HPDIs)))

@@

  return(ggplot(visContrastTreatmentIC, aes(x = IC, y = contrast, fill = group ))+
           geom_line(se=FALSE)+
    geom_ribbon(mapping = aes(ymin = cLow, ymax = cHigh), alpha = .3)+
    ylim(c(-2,2)) +theme_tufte()+geom_hline(yintercept = 0, lty =2, size = 0.1))
}

visContrastsICJoint <- ggarrange(
visContrastsIC(ADS = 2, CBS = -2)+removeX+ 
               theme(legend.position = c(0.3, 0.9),
                     legend.key.size = unit(.3, 'cm'),
                     legend.key.height = unit(.3, 'cm'),
                     legend.key.width = unit(.3, 'cm'),
                     legend.title= element_blank())+
  ggtitle("CBS = -2")+
  ylab("ADS = 2"),
    visContrastsIC(ADS = 2, CBS = 0)+removeY+removeX+ scale_fill_discrete(guide=FALSE)+
  ggtitle("CBS = 0"),
    visContrastsIC(ADS = 2, CBS = 2)+removeY+removeX+ggtitle("CBS = 2")+ scale_fill_discrete(guide=FALSE),
visContrastsIC(ADS = 0, CBS = -2)+removeX+ scale_fill_discrete(guide=FALSE)+
  ylab("ADS = 0"),
    visContrastsIC(ADS = 0, CBS = 0)+removeY+removeX+ 
  scale_fill_discrete(guide=FALSE),
    visContrastsIC(ADS = 0, CBS = 2)+removeY+removeX+ scale_fill_discrete(guide=FALSE),  
visContrastsIC(ADS = -2, CBS = -2)+ scale_fill_discrete(guide=FALSE)+
  ylab("ADS = -2"),
    visContrastsIC(ADS = -2, CBS = 0)+removeY+ scale_fill_discrete(guide=FALSE),
    visContrastsIC(ADS = -2, CBS = 0)+removeY+ scale_fill_discrete(guide=FALSE), 
ncol = 3, nrow = 3
)

@@

# Direct effect

Models for the evaluation of direct effect need to close the indirect causal path from the treatment variables to the output, and they do so by conditioning on \textsf{CAS}. Again, we face model selection. We repeat all the model structures from the previous section, except for now, each model is extended with \textsf{CAS} as a predictor. This time the model structure we called \textsf{tooFar} turns out to do better. We also consider extending it with interactions between \textsf{CAS} and \textsf{IT} and \textsf{IC} (jointly and separately), but this results in no further improvement.

\vspace{1mm}
\footnotesize
```{r comparisonDirectModels,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
nullDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bCAS * CAS,
    a ~ dnorm (0,0.3),
    bCAS ~ dnorm(0,0.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries  
)

ADSDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <-  a + bADS * ADS+ bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,0.3),
    bCAS ~ dnorm(0,0.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSICDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <-  a + bADS * ADS+ bIC * IC+ bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,0.3),
    bIC ~ dnorm(0,0.3),
    bCAS ~ dnorm(0,0.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ITDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <-  bIT[groupID] + bCAS * CAS,
    bIT[groupID] ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS * ADS +  bIT[groupID]+ bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bIT[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITICDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS * ADS +  bIT[groupID] + bIC * IC+ bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITIC_ADSICDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS * ADS +  bIT[groupID] + bIC * IC + bADSIC * ADS * IC+ bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bADSIC ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITIC_ADSIC_ADSITDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC * IC + bADSIC * ADS * IC+ bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSIT_ADSITDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    #bADSIC ~ dnorm(0,.5),
    bIT[groupID] ~ dnorm(0,.3),
    #bIC ~ dnorm(0,.5),
    bCAS ~ dnorm(0,0.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITIC_ADSIT_ITIC_ADSICDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC +
      bADSIC * ADS * IC + bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

ADSITICCBS_ITIC_ADSICDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
      bADSIC * ADS * IC+ bCBS *CBS + bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCBS ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

FinalDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
      bADSIC * ADS * IC+ bCBS[groupID] *CBS + bCAS * CAS,
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bCBS[groupID] ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

tooFarDirect <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
      bADSIC * ADS * IC+ bCBS[groupID] *CBS + bCBSIC * CBS * IC + bCAS * CAS, 
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bCBS[groupID] ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    bCBSIC ~ dnorm(0, .3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

tooFarDirect_CASIT <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
      bADSIC * ADS * IC+ bCBS[groupID] *CBS + bCBSIC * CBS * IC + bCAS[groupID] * CAS, 
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCAS[groupID] ~ dnorm(0,0.3),
    bCBS[groupID] ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    bCBSIC ~ dnorm(0, .3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

tooFarDirect_CASIC <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
      bADSIC * ADS * IC+ bCBS[groupID] *CBS + bCBSIC * CBS * IC + bCAS * CAS + bCASIC * CAS * IC, 
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCAS ~ dnorm(0,0.3),
    bCASIC ~ dnorm(0,0.3),
    bCBS[groupID] ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    bCBSIC ~ dnorm(0, .3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

tooFarDirect_CASIC_CASIT <- quap(
  alist(
    AdiffS ~ dnorm( mu, sigma ),
    mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
      bADSIC * ADS * IC+ bCBS[groupID] *CBS + bCBSIC * CBS * IC + bCAS[groupID] * CAS + bCASIC * CAS * IC, 
    a ~ dnorm (0,0.3),
    bADS[groupID] ~ dnorm(0,.3),
    bADSIC ~ dnorm(0,.3),
    bCAS[groupID] ~ dnorm(0,0.3),
    bCASIC ~ dnorm(0,0.3),
    bCBS[groupID] ~ dnorm(0,.3),
    bIT[groupID] ~ dnorm(0,.3),
    bIC[groupID] ~ dnorm(0,.3),
    bCBSIC ~ dnorm(0, .3),
    sigma  ~ dexp(1)
  ), 
  data = summaries
)

comparisonDirect<- compare(nullDirect,ADSDirect,ADSICDirect,ITDirect,ADSITDirect,ADSITICDirect,ADSITIC_ADSICDirect, ADSITIC_ADSIC_ADSITDirect,ADSIT_ADSITDirect,ADSITIC_ADSIT_ITIC_ADSICDirect,
                     ADSITICCBS_ITIC_ADSICDirect,FinalDirect, tooFarDirect,tooFarDirect_CASIT,
                     tooFarDirect_CASIC,tooFarDirect_CASIC_CASIT)

mykable(round(data.frame(comparisonDirect),3))%>% kable_styling(latex_options = c("striped","scale_down"),font_size = 9)

plot(comparisonDirect)

Let's build a Hamiltonian Monte Carlo with the same formula:

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}

# tooFarDirectHMC <- ulam(
#   alist(
#     AdiffS ~ dnorm( mu, sigma ),
#     mu <- a + bADS[groupID] * ADS +  bIT[groupID] + bIC[groupID] * IC + 
#       bADSIC * ADS * IC+ bCBS[groupID] *CBS + bCBSIC * CBS * IC + bCAS * CAS, 
#     a ~ dnorm (0,0.3),
#     bADS[groupID] ~ dnorm(0,.3),
#     bADSIC ~ dnorm(0,.3),
#     bCAS ~ dnorm(0,0.3),
#     bCBS[groupID] ~ dnorm(0,.3),
#     bIT[groupID] ~ dnorm(0,.3),
#     bIC[groupID] ~ dnorm(0,.3),
#     bCBSIC ~ dnorm(0, .3),
#     sigma  ~ dexp(1)
#   ), 
#   data = summaries
# )

#saveRDS(tooFarDirectHMC, file = "models/tooFarDirectHMC.rds")

tooFarDirectHMC <- readRDS(file = "models/tooFarDirectHMC.rds")
```
\normalsize

For a big picture, let's look at predicted direct effect by group and user activity profile.

visGroupDirect <- function (model, ADS, CBS, CAS, xmin =2, ymax = -3)
{
  groupID <- 1:3
  IC <- 5 
  data <- expand.grid(ADS = ADS,groupID = groupID, CBS = CBS, CAS = CAS, IC =  IC)
  posterior <- extract.samples(model, n = 1e5)
  mu <- link( model, data=data ) 
  colnames(mu) <- levels(summaries$group)
  muLong <- melt(mu)
  colnames(muLong) <- c("id", "group", "AdiffS")
  means <-  round(apply(mu , 2 , mean ), 2)
  mu_HPDI <- round(apply( mu , 2 , HPDI ),2)
  means <- as.data.frame(means)
  means$group <- rownames(means)
  rownames(means) <- NULL
  meansDisp <- cbind(means,t(as.data.frame(mu_HPDI)))
  meansDisp <- meansDisp[,c(1,3,4)]

  plot <- ggplot(muLong)+geom_violin(aes(x = group, y = AdiffS), alpha = 0.2)+
    xlab("")+
    labs(title = paste("ADS=", ADS, ", CBS=",  CBS, ", CAS=", CAS,  sep = ""))+
    theme_tufte()+ylim(c(-4,4))
  #+   annotation_custom(tableGrob(meansDisp), xmin=xmin,  ymax=ymax)
  return(plot)
}

visGroupDirect2_2_2 <-  visGroupDirect(model = tooFarDirectHMC, ADS = 2,CBS = -2, CAS = -2)
visGroupDirect200 <-  visGroupDirect(model = tooFarDirectHMC, ADS = 2,CBS = 0, CAS = 0)
visGroupDirect222 <- visGroupDirect(model = tooFarDirectHMC, ADS = 2,CBS = 2, CAS = 2)

visGroupDirect0_2_2 <-  visGroupDirect(model = tooFarDirectHMC, ADS = 0,CBS = -2, CAS = -2)
visGroupDirect000 <-  visGroupDirect(model = tooFarDirectHMC, ADS = 0,CBS = 0, CAS = 0)
visGroupDirect022 <- visGroupDirect(model = tooFarDirectHMC, ADS = 0,CBS = 2, CAS = 2)

visGroupDirect_2_2_2 <-  visGroupDirect(model = tooFarDirectHMC, ADS = -2,CBS = -2, CAS = -2)
visGroupDirect_200 <-  visGroupDirect(model = tooFarDirectHMC, ADS = -2,CBS = 0, CAS = 0)
visGroupDirect_222 <- visGroupDirect(model = tooFarDirectHMC, ADS = -2,CBS = 2, CAS = 2)

visGroupDirectJoint <- ggarrange(visGroupDirect2_2_2+removeX +ylab("ADS = 2")+ggtitle("CBS = CAS = -2"),
                                 visGroupDirect200+removeX+removeY+ ggtitle("CBS = CAS = 0"),
                                 visGroupDirect222+removeX+removeY+ ggtitle("CBS = CAS = 2") ,
                        visGroupDirect0_2_2+removeX + ggtitle("")+ylab("ADS = 0"),
                                visGroupDirect000+removeX+removeY+ ggtitle(""),
                                visGroupDirect022+removeX+removeY+ ggtitle(""),
                        visGroupDirect_2_2_2+removeX+ylab("ADS = -2")+ ggtitle(""),
                                  visGroupDirect_200+removeX+removeY+ ggtitle(""),
                        visGroupDirect_222+removeX+removeY+ ggtitle("")
                                 ,ncol =3, nrow = 3)

visGroupDirectJoint2 <- annotate_figure(visGroupDirectJoint, 
                                  top = text_grob("(range restricted to (-4,4), IC at the rounded mean = 5)",
                                                  size = 10))
visGroupJoint3 <- annotate_figure(visGroupDirectJoint2, 
                                  top = text_grob("Predicted direct effect of treatment group by activity profile (standardized)",
                                                  size = 12))

Now, predicted direct effect of five interventions vs. \textsf{CBS}, for three user activity profiles.

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
visContrastsCBSDirect <- function(model = FinalHMC, ADS = ADS , CAS= CAS, IC =  5,
                            CBS = seq(-3,3,by  = 0.1)){
  groupID <- 1:3
  data <- expand.grid(ADS, groupID, CAS, IC , CBS)
  colnames(data) <- c("ADS", "groupID", "CAS", "IC", "CBS")
  posterior <- extract.samples(model, n = 1e5)
  link( model, data=data ) 
  mu <- link( model, data=data )

  means <-  round(apply(mu , 2 , mean ), 4)

  HPDIs <- round(apply( mu , 2 , HPDI ),4)
  visContrast <- cbind(data,means,t(as.data.frame(HPDIs)))

  colnames(visContrast)[c(7,8)] <- c("low", "high")

  contrast <- numeric(nrow(visContrast))
  cLow <- numeric(nrow(visContrast))
  cHigh <- numeric(nrow(visContrast))
  for(i in threes){
    contrast[i] <- visContrast$means[i] - visContrast$means[i-2]  
  }
  for(i in twos){
    contrast[i] <- visContrast$means[i] - visContrast$means[i-1]  
  }
  visContrast$contrast <- contrast
  visContrast$shift <-  visContrast$contrast - visContrast$means
  for(i in ones){
    visContrast$shift[i] <- 0
  }
  visContrast$cLow <- visContrast$low + visContrast$shift
  visContrast$cHigh <- visContrast$high + visContrast$shift

  return(ggplot(visContrastTreatment, aes(x = CBS, y = contrast, fill = group ))+
           geom_line(se = FALSE)+
           geom_ribbon(mapping = 
                         aes(ymin = cLow, ymax = cHigh),  
                       alpha = .3)+
           theme_tufte()+ylim(c(-3.5,3.5)))
}

visContrastDirect <-  ggarrange(visContrastsCBSDirect(model = tooFarDirect,ADS = -2, CAS = -2, IC = 5)+ggtitle("ADS = CAS = -2")+ scale_fill_discrete(guide=FALSE),
          visContrastsCBSDirect(model = tooFarDirect,ADS = 0, CAS = 0, IC = 5) +ggtitle("ADS = CAS = 0")+ scale_fill_discrete(guide=FALSE)+removeY,
          visContrastsCBSDirect(model = tooFarDirect,ADS = 2, CAS = 2, IC = 5) +ggtitle("ADS = CAS = 2")+ scale_fill_discrete(guide=FALSE)+removeY
          ,ncol =3)
